---
title: "sensitive-period-modeling"
author: "Ben"
date: "6/2/2018"
output: html_document
---

```{r libraries, message=FALSE, warning=FALSE}
library(rstan)
library(tidyverse)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```


```{r data}
df <- read.csv('osf-resources/data.csv')
```

# Subset analysis
```{r subset}
n_samples <- 1000
n_sample_ids <- sample(df$id, size=n_samples)
df_subset <- df %>%
  filter(id %in% n_sample_ids)
```

```{r reduce-cols}
cols_demo <- c('id', 'gender', 'age', 'primelangs', 'education', 'tests', 'Eng_start', 'Eng_country_yrs', 'house_Eng', 'currcountry')
# Melt question cols
df_subset_melt <- df_subset %>% 
  select(cols_demo, contains('q')) %>%
  gather(question, answer, contains('q'))
```

```{r plot-proportion-correct}
df_subset_melt %>%
  group_by(question) %>%
  summarise(mean_answer=mean(answer)) %>%
  mutate(multi_part_q=grepl('_', question)) %>%
  ggplot(aes(x=question, y=mean_answer, fill=multi_part_q)) +
    geom_bar(stat='identity', position='dodge') +
    coord_cartesian(ylim=c(0.5,1.0)) +
    theme_classic()
```


```{r}
M <- df_subset %>% select(id, contains('q'))
```


## Modeling

  * `ability` parameter for each participants
  * `difficulty` parameter for each question


Probability a participant correctly answer a question
$$\sigma^{-1}(p_{qs}) = \mu_0 + \eta_q + \eta_s$$

The log odds of the probabiliyt of a correct response for question $q$ from participant $s$ is modeled by some overall mean $\mu_0$ along with contributes for the difficulty of the question $\eta_q$ and latent ability of participant $\eta_s$.

We assume:

$$\sigma^{-1}(p_{i}) = \mu_0 + \eta_q + \eta_s$$
$$\eta_q \sim N(0, \sigma^2_{Q}$$
$$\eta_s \sim N(0, \sigma^2_{S}$$
Data must be in long-form.
```{r}
# Check for NAs
M_melt <- M %>%
  gather(question, answer, contains('q')) %>%
  filter(question %in% c("q1", "q2", "q3"))
M_melt %>% 
  group_by(answer) %>%
  summarise(n=n())
```

```{r}
stan_model <- "
data {
  int<lower=1> I;               // # n_questions
  int<lower=1> J;               // # n_participants
  int<lower=1> N;               // # total observations
  int<lower=1, upper=I> ii[N];  // question for n
  int<lower=1, upper=J> jj[N];  // person for n
  int<lower=0, upper=1> y[N];   // correctness for n
}
parameters {
  vector<lower=0>[I] alpha;     // discrimination for item i
  vector[I] beta;               // difficulty for item i
  vector[J] theta;              // ability for person j
}
model {
  vector[N] eta;
  alpha ~ lognormal(0.5,1);
  beta ~ normal(0,10);
  theta ~ normal(0,1);
  for (n in 1:N)
    eta[n] = alpha[ii[n]] * (theta[jj[n]] - beta[ii[n]]);
  y ~ bernoulli_logit(eta); 
}
"

```

Note:
  * Data cannot contains missing values (stan doesn't like)
  * Identifiers for students and items must be consecutive intevers starting with 1
  
  
```{r stan-data-setup}
M_melt$participant_id <- 1:nrow(M_melt)
unique_questions <- unique(M_melt$question)
question_ids <- 1:length(unique_questions)
df_question_map <- data.frame(question=unique_questions, question_id=question_ids)
question2id <- function(question_) {
  subset(df_question_map, question==question_)$question_id
}
M_melt$item_id <- purrr::map_int(M_melt$question, question2id)
M_melt <- M_melt %>%
  rename(response=answer) %>%
  select(participant_id, item_id, response)

stan_data <- list(
  I=max(M_melt$item_id),
  J=max(M_melt$participant_id),
  N=nrow(M_melt),
  ii=M_melt$item_id,
  jj=M_melt$participant_id,
  y=M_melt$response)

```



```{r}
fit <- stan(model_code=stan_model, data=stan_data,model_name='sensitive_period_2pl', iter=500, chains=4)
```

# Model checking

Get params
```{r params}
print(fit, pars=c('alpha', 'beta'))

all_estimates <- summary(fit)[[1]]
theta_estimates <- summary(fit, pars='theta')[[1]]
```

Convergence
```{r convergence}
stan_rhat(fit, 'theta')
```








